---
title: "Project 1"
author: "Jingjing Li"
date: "6/13/2021"
output: github_document
    toc: true  
    toc_depth: 1  
    toc_float: false  
    
---
# Contents
## Introduction  
This project aims to create a vignette for reading and summarizing data from the National Hockey League’s (NHL) API.  


## Link to repo  
## Required packages  
 * httr  
 * jsonlite  
 
## Functions  
### NHL records API 
#### Load packages
```{r, warning=FALSE, message=FALSE}
# load package to get data from API link
library(httr)
library(jsonlite)
library(dplyr)
```

#### franchise function
```{r}
# Construct a function for API data acquisition

franchise <- function (id,name){
  base_url <- "https://records.nhl.com/site/api/franchise"
  get_url <- GET (base_url)
  record_txt <- content (get_url, "text", encoding = "UTF-8")
  record_json <- fromJSON(record_txt, flatten=TRUE)
  Data_df <- data.frame( record_json)
  if (id==0 & name == "full") {
   return(Data_df) 
  }
   else{record <- filter(Data_df, data.id==id | data.fullName==name)
   return(record)
   }
}
record_franchise1 <- franchise (2, "Montreal Wanderers")
record_franchise1
record_franchise2 <- franchise (0,"full")
record_franchise2
```
```{r}
total <- function (id, name){
  base_url <- "https://records.nhl.com/site/api"
  full_url <- paste0 (base_url, "/franchise-team-totals")
  get_url <- GET (full_url)
  record_txt <- content (get_url, "text", encoding = "UTF-8")
  record_json <- fromJSON(record_txt, flatten=TRUE)
  Data_df <- data.frame( record_json)
   if (id==0 & name == "full") {
   return(Data_df) 
  }
   else{record <- filter(Data_df, data.franchiseId==id | data.teamName==name )
   }
  
  return(record)
  }
record_total1 <- total (2, "Montreal Wanderers")
record_total1
record_total2 <- total (0, "full")
record_total2
```

#### season function
```{r}
season <- function (id){
  base_url <- "https://records.nhl.com/site/api/franchise-season-records?cayenneExp=franchiseId"
  if (id==0 ) {
   get_url <- GET (base_url) 
   record_txt <- content (get_url , "text", encoding = "UTF-8")
   record_json <- fromJSON(record_txt, flatten=TRUE)
   Data_df <- data.frame(record_json)
    return( list(base_url,Data_df))
  }
    else {
      full_url  <-  paste0(base_url,"=",id )	
      get_url <- GET (full_url)
      record_txt <- content (get_url , "text", encoding = "UTF-8")
      record_json <- fromJSON(record_txt, flatten=TRUE)
      Data_df <- data.frame(record_json)
      return( list(full_url,Data_df))
    }

 
}
record_season <- season (9)
record_season       
record_season <- season (0)
record_season       
```
#### goalie function
```{r}

goalie<- function (id){
  base_url <- "https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId"
   if (id==0 ) {
  get_url <- GET (base_url) 
   record_txt <- content (get_url , "text", encoding = "UTF-8")
   record_json <- fromJSON(record_txt, flatten=TRUE)
   Data_df <- data.frame(record_json)
    return( list(base_url,Data_df))
  }
    else {
      full_url  <-  paste0(base_url,"=",id )	
      get_url <- GET (full_url)
      record_txt <- content (get_url , "text", encoding = "UTF-8")
      record_json <- fromJSON(record_txt, flatten=TRUE)
      Data_df <- data.frame(record_json)
      return( list(full_url,Data_df))
    }
}
record_goalie1<- goalie (6)
record_goalie1
str(record_goalie1)
record_goalie2 <- goalie (0)
record_goalie2
 
```
skater function
```{r}
skater <- function (id,...){
  base_url <- "https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId"
  if (id==0 ) {
get_url <- GET (base_url) 
   record_txt <- content (get_url , "text", encoding = "UTF-8")
   record_json <- fromJSON(record_txt, flatten=TRUE)
   Data_df <- data.frame(record_json)
    return( list(base_url,Data_df))
  }
    else {
      full_url  <-  paste0(base_url,"=",id )	
      get_url <- GET (full_url)
      record_txt <- content (get_url , "text", encoding = "UTF-8")
      record_json <- fromJSON(record_txt, flatten=TRUE)
      Data_df <- data.frame(record_json)
      return( list(full_url,Data_df))
    }
}
record_skater1 <- skater( 6 )
record_skater1
str(record_skater1)
record_skater2 <- skater( 0 )
record_skater2
```
#### detail function
```{r}
detail<- function (id,...){
  base_url <- "https://records.nhl.com/site/api/franchise-detail?cayenneExp=mostRecentTeamId"
  if (id==0 ) {
get_url <- GET (base_url) 
   record_txt <- content (get_url , "text", encoding = "UTF-8")
   record_json <- fromJSON(record_txt, flatten=TRUE)
   Data_df <- data.frame(record_json)
    return( list(base_url,Data_df))
  }
    else {
      full_url  <-  paste0(base_url,"=",id )	
      get_url <- GET (full_url)
      record_txt <- content (get_url , "text", encoding = "UTF-8")
      record_json <- fromJSON(record_txt, flatten=TRUE)
      Data_df <- data.frame(record_json)
      return( list(full_url,Data_df))
  }
}
record_detail <- detail( 2 )
record_detail
record_detail <- detail( 0 )
record_detail
```


### NHL stats API

```{r}
stats <- function(id, ... ) {
  base_url2 <- "https://statsapi.web.nhl.com/api/v1/teams/" 
  if (is.numeric (id)) {
    apiurl <- paste0 (base_url2, id, "/?expand=team.stats")
  }
  if (!is.numeric (id)) {
      apiurl<- paste0 (base_url2, "?expand=team.stats")
    }
  team <- GET (apiurl)
  team_txt <- content (team,"text", encoding="UTF-8")
  stats_json <- fromJSON(team_txt, flatten=TRUE)
  team_df <- data.frame(stats_json) 
  return(team_df)
} 
   stats_data <- stats (2)
   stats_data
   stats_data2 <- stats ("teams")
   stats_data2
```
## wrapper function
```{r}
wrapper <- function (modifier, id,name,...) {
  if (modifier == "franchise") {
    franchise_Data <- franchise (id,name)
    franchise_Data
  }
    else if (modifier == "totals"){
     total_Data <- total (id,name)
     total_Data
    }
      else if (modifier == "season") {
       season_Data <- season(id)
       season_Data
      }
        else if (modifier == "goalie"){
         goalie_Data <- goalie (id)
         goalie_Data
        }
          else if (modifier == "skater"){
            skater_Data <- skater (id)
            skater_Data
          }
            else if (modifier == "detail") {
              detail_Data <-  detail (id)
              detail_Data
            }
              else if (modifier == "stats"){
                stas_Data <- stats (id)
                stas_Data
              }
}
result1 <- wrapper ("franchise", 2, "Montreal Wanderers")
result1
result2 <- wrapper ("stats", 2)
result2
result3 <- wrapper ("skater",0)
result3
```
  
## Exploratory Data analysis 

```{r}
goalie_data <- wrapper ("goalie", 6)
goalie_data <- data.frame(goalie_data) %>% group_by (data.franchiseId, data.franchiseName  )
skater_data <- wrapper ("skater", 6)
skater_data <- data.frame (skater_data)
franchise_data <- wrapper ("franchise", 2, "Montreal Wanderers")
franchise_data
total_Data <- wrapper ("totals", 2, "Montreal Wanderers")
total_Data
gp_sk_comb <- inner_join(goalie_data, skater_data, by=(c("data.franchiseId","data.franchiseName"))

```

### Data from two endpoints
### New variables
### contingency tables
### Five plots
#### bar plot
#### histogram
#### box plot
#### scatter plot
